---
title: "assignment_2"
format: html
editor: visual
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo       = TRUE, 
                      fig.height = 3, 
                      fig.width  = 6,
                      fig.align  = "center")
ggplot2::theme_set(ggplot2::theme_bw())
```

```{r}
#| message: false
library(tidyverse)

library(lubridate)
```

# Question 1:

Write a function `pos_na()` that takes two vectors of equal length and returns the positions where **both** vectors contain `NA`.\
If the vectors do **not** have the same length, the function should return the message:

> `"The vectors are not the same length."`

Use the following pairs of vectors to test your function:

1.  `c(1, NA, 2)` and `c(NA, NA)`
2.  `c(NA, NA, 2)` and `c(NA, NA, 2)`
3.  `c(NA, 5, NA)` and `c(NA, NA, NA)`
4.  `c(NA, NA, NA, 2, NA, 4, NA)` and `c(NA, NA, 2, 4, 4, NA, NA)`
5.  `c(1, NA, NA, 2, NA, 4, NA)` and `c(NA, -5, 2, 4, 4, NA, 11)`

Here’s a polished, professional way to label those two parts of the assignment. It reads cleanly and sets expectations without adding extra wording.

------------------------------------------------------------------------

### (1i) Using an `if` statement

Write a version of `pos_na()` that checks whether the two input vectors have the same length using an `if` statement.

-   If they do, return the positions where both vectors contain `NA`.

-   If they do not, return the message:

> `"The vectors are not the same length."`

```{r}
v1 = c(1, NA, 2) 
v2 = c(NA, NA, 3)

pos_na <- function(vector1, vector2) {
  if (length(vector1) == length(vector2)) {
    positon_NA_1 <- which(is.na(vector1))  
    position_NA_2 <- which(is.na(vector2))
    return(list(positon_NA_1, position_NA_2))
  }
  else {
    return("The vectors are not the same length.")
  }
}
  
pos_na(v1, v2)



```

```{r}

v3 = c(1, NA, 2) 
v4 = c(NA, NA)
pos_na(v3, v4)


```

### (1ii) Without using an `if` statement

Write a second version of `pos_na()` that performs the same task **without** using an `if` statement.

-   You may use any other valid R mechanism (e.g., `stop()`, `stopifnot()`, or logical short‑circuiting) to handle the case where the vectors differ in length.

```{r}

```

```{r}

pos_na_2 <- function(vector1, vector2) {
  stopifnot(length(vector1) == length(vector2))
  return(list(which(is.na(vector1)), which(is.na(vector2))))
  }

pos_na_2(v3, v4) 

```

```{r}

v5 = c(NA, NA, NA, 2, NA, 4, NA) 
v6 = c(NA, NA, 2, 4, 4, NA, NA)
pos_na_2(v5, v6)


```

# Question 2

Load the `wmata_ridership` data frame into R from <https://dcgerard.github.io/stat_412_612/data/wmata_ridership.csv>. For each month, calculate the proportion of rides made on a given day of the month. Then make box plots of ridership proportions by day of the week. But exclude any days from 2004 and 2005.

```{r}

wamata_ridership <- read_csv("https://dcgerard.github.io/stat_412_612/data/wmata_ridership.csv")

tail(wamata_ridership)

```

```{r}

wamata_ridership <- wamata_ridership |> 
      mutate(Date = ymd(Date),
             Year = year(Date),
             Month = month(Date),
             DayOfWeek = wday(Date, label = TRUE, abbr = TRUE),
             DayofMonth = day(Date)) 


wamata <- wamata_ridership |>
  filter(!(Year %in% c(2004,2005)))

monthly_totals <- wamata |>
  group_by(Year, Month) |>
  summarise(monthly_rides = sum(Total), .groups = "drop")

wamata_prob <- wamata |>
  left_join(monthly_totals, by = c("Year", "Month")) |>
  mutate(proportions = Total / monthly_rides)

ggplot(wamata_prob, aes(x = DayOfWeek, y = proportions)) + geom_boxplot()  + 
  labs(title = "Boxplot of Wamata Ridership Proportions",
       y = "Proportion of Monthly Rides",
       x = "Day of the Week") 
```

```{r}

# for month for date calculate date/ total month 
# mutate new column for proportion 
# boxplot of new column 



```

# Question 3:

Write R code that extracts the elements "Bears", "Dolphins", and "Bengals" from the vector `V` shown below, and display the result.

```{r}
V <- c("Bears", "Lions", "Dolphins", "Eagles", "Bengals")
```

## (3i) Subsetting by index

```{r}

V[1]
V[2]
V[3]
V[4]
V[5]
```

## (3ii) Subsetting using a `for` loop

```{r}

for (i in V) {
  print(i)
}

```

## (3iii)Returning all values in a single vector (one line of code)

### (a) Using positive indices

```{r}

V[1:5]

```

### (b) Using negative indices

```{r}

V[-6]

```

### (c) Using logical indexing

```{r}

V[]

```

### (d) Attempting to subset by names

```{r}

names(V) <- c("a", "b", "c", "d", "e")

V["a"]
V["b"]
V["c"]
V["d"]
V["e"]

```

# Question 4:

You are given a character vector containing messy employee records. Each entry includes:

-   an employee name (string),
-   a hire date in inconsistent formats,
-   and a department code that should be treated as a factor.

```{r}
records <- c(
  "Steve McQueen | 2020-01-15 | HR",
  "B. Smith | 15/02/2021 | FIN",
  "Carlos M | March 3, 2019 | IT",
  "D. Lee | 2018/07/30 | HR",
  "Alain Delon | 04-12-2020 | MKT"
)
records
```

1.  **Parse each record** into three separate variables:

    -   `name`\
    -   `hire_date`\
    -   `dept`

```{r}

split_records <- str_split(records, "\\|")

name <- c(split_records[[1]][1], split_records[[2]][1], split_records[[3]][1], split_records[[4]][1], split_records[[5]][1])

hire_date <- c(split_records[[1]][2], split_records[[2]][2], split_records[[3]][2], split_records[[4]][2], split_records[[5]][2])

dept <- c(split_records[[1]][3], split_records[[2]][3], split_records[[3]][3], split_records[[4]][3], split_records[[5]][3])

dept
```

2.  **Convert `hire_date` into a proper `Date` object**, correctly handling all of the different date formats.

```{r}

hire_date <- c(parse_date(hire_date[1], format = "%Y-%m-%d"), parse_date(hire_date[2], format = "%d/%m/%Y"), mdy(hire_date[3]), parse_date(hire_date[4], format = "%Y/%m/%d"), parse_date(hire_date[5], format = "%m-%d-%Y"))

hire_date
```

3.  **Convert `dept` into an ordered factor** with the levels:\
    `HR`, `FIN`, `IT`, `MKT`.

```{r}

dept <- str_trim(dept)

dept <- factor(dept, levels = c("HR", "FIN", "IT", "MKT"), ordered = TRUE)

dept


```

4.  **Construct a clean data frame** containing the parsed and converted variables.

```{r}

hiring <- tibble(name, hire_date, dept)

hiring
```

1.  **Add a new variable** called `years_worked` that reports the number of full years each employee has worked as of today.

1.  **Return the final cleaned data frame.**
